Low Complexity Adaptive Sonar Imaging

### 2016-02-17 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2016:02:17	16:44:20	(on.17.feb.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image by adapting the array's weights, MVDR seeks an optimal solution by computing the spatial statistics of the wavefield, while LCA simply selects the best window out of a predefined set.

We show that a robust MVDR implementation typically creates windows spanning between the rectangular and Hamming. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the -3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS 1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

\IEEEPARstart{T}{he} best adaptive sonars are found in nature. Bats, for instance, use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. Active human made sonars have little to no adaptability on transmit, but several methods have been investigated to adaptively post process the received data. Perhaps the best studied method for reconstructing a sonar image is the minimum variance distortion-less response (MVDR) beamformer~\cite{Capon1969}. It adapts the sonar array's spatial response to minimize the influence of noise and interference in the final image.

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% \begin{itemize}
% \item Dynamic weights $\Rightarrow$ greater contrast/resolution potential
% \item But:
% \begin{itemize}
% \item Active system $\Rightarrow$ noise and signal correlated $\Rightarrow$ adaptive beamformers no longer robust.
% \item Subarray averaging decorrelate the noise, but in the process we sacrifice resolution.
% \item Computationally complex O($M^3$) due to inversion of a spatial covariance matrix. LCA is only O($MW$). Even if Capon subspace/beamspace versions exist, they remain more complex than the LCA, and is not as ideally suited for implementation of SIMD hardware (GPUs, DSPs, ...).
% \end{itemize}
% \item LCA is inherently robust, is ``parameter free'' once setup with a well designed window set, and can often be implemented with minor modifications to already existing DAS beamformers.
% 
% The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.
% 
% While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output. 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR achieves this by computing optimal array weights from the spatial statistics of the wavefield, while LCA selects the best performing weights out of a predefined set.

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the -3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS 1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

% The promise is already demonstrated in nature.

The sonar systems of the future

Future sonar systems

Great promise in adaptivity

The greatest untapped

The sonars of the future might benefit greatly from adaptivity.

Future sonars will likely be

\IEEEPARstart{A}{daptive} sonars holds great promise. This promise is allured to in nature by for instance bats, which use a high frequency sonar to detect, identify and track prey with amazing precision in caves amongst a myriad of other bats. No man made sonar can currenly match this feat, because it is hard to determine what the correct action is for an arbitrary situation, and too slow to process all the information required to typically slow to process all the information needed to act correctly the situation in time, and non trivial to determine exactly how it should adapt. In current active sonars there is little to no adaptability on transmit, several methods have been investigated to adaptively post process the received data. Perhaps the best studied method for reconstructing a sonar image is the minimum variance distortion-less response (MVDR) beamformer~\cite{Capon1969}. It adapts the sonar array's spatial response to minimize the influence of noise and interference in the final image.

### 2016-02-17 LCAforActiveSonarImaging/submission/submit1/answer/answer.tex
la2pdf commit - 2016:02:17	16:44:20	(on.17.feb.)

\q{The paper is well written and addresses an actual problem. The solution however, is barely innovative. Still the results are interesting enough for publication. Some minors details.} \\\\
\ans{We appreciate the reviewer's }

@@

\begin{enumerate}
% \emph{General}
\hi{Introduction}
\qi{The authors show no evidence that bats use adaptive processing, remove this example.}
\ai{The reviewer makes a good point, this may not have been obvious. We have provided a reference for the adaptive processing used by bats, but would prefer to keep the example if possible.}
%
\qi{It is not clear whether LCA is really an adaptive method that adapts weights on the fly or a method that a posteriori decides which window shows best results.}
\ai{We appreciate having this pointed out. We are using a rather loose definition of adaptivity that includes any kind of dynamic and data dependent behaviour, hence feel that both beamformers fall in that category. However, as the reviewer pointed out, the method . }
%
\qi{The use of non-SI units like cm is discouraged.}
\ai{We agree that non-SI units should be avoided, however had the belief that cm was a valid unit? In the IEEE Styles Manual this unit is used several times as an example: 
\url{http://www.ieee.org/publications\_standards/publications/journmag/online\_style\_manual-10292015.pdf}} %(on page 24 paragraph ``The En, Em, or Two-Em Dash'', and several times on page 25 paragraph ``Math'' bullet point 16. Should we be adhering to a different style guide?}
%
\end{enumerate}

@@

\ans{Thank you for the thorough review and very constructive remarks. We believe in this philosophy and will make changes accordingly.}

>>>>

\q{The paper is well written and addresses an actual problem. The solution however, is barely innovative. Still the results are interesting enough for publication. Some minors details.} \\\\
\ans{We appreciate the reviewer's view and would like to thank him for this time and patience reviewing this article.}

@@

\begin{enumerate}
% \emph{General}
\hi{Introduction}
\qi{The authors show no evidence that bats use adaptive processing, remove this example.}
\ai{The reviewer makes a good point, this may not have been obvious. We have provided a reference for the adaptive processing used by bats, but would prefer to keep the example if possible.}
%
\qi{It is not clear whether LCA is really an adaptive method that adapts weights on the fly or a method that a posteriori decides which window shows best results.}
\ai{We appreciate having this pointed out. We are using a rather loose definition of adaptivity that includes any kind of dynamic and data dependent behaviour, hence feel that both beamformers fall in that category. However, as the reviewer pointed out, the method . }
%
\qi{The use of non-SI units like cm is discouraged.}
\ai{We agree that non-SI units should be avoided, however had the belief that cm being a prefixed SI unit was valid? In the IEEE Styles Manual this unit is used several times as an example: 
\url{http://www.ieee.org/publications_standards/publications/journmag/online_style_manual-10292015.pdf} (on page 24 paragraph ``The En, Em, or Two-Em Dash'', and several times on page 25 paragraph ``Math'' bullet point 16. Will leave this as is for now pending an answer for this.}
%
\end{enumerate}

@@

\ans{We thank the reviewer for the thorough review and very constructive remarks. We believe in this philosophy and will make changes accordingly.}
